---
title: "Programming in Base R"
format: html
editor: visual
---

## Task 1: Basic Vector Practice 

Suppose we have data from a medical experiment on blood pressure. 


Assign objects pre-measurements and post-measurements 
```{r}
pre_bp <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
pre_bp

post_bp <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
post_bp
```

Assign names to the vector elements using paste() function
```{r}
patient <- paste("Subject", 1:20, sep = "_")
```

Calculate the change in blood pressure for each patient by subtracting post-treatment measurements from pre-treatment measurements.Create Vector for this object.
```{r}
diff_bp <- c(pre_bp - post_bp)
diff_bp
```
Calculate the average decrease in blood pressure across all patients using mean() function 
```{r}
mean(diff_bp)
```

Determine which patients experienced a decrease in blood pressure after treatment using which() function 

return the indices and names associated with this type of change
```{r}
which (diff_bp > 0)
```

Create a vector for patients that experienced a decrease in blood pressure



## Task 2: Basic Data Frame Practice 

Create a data frame with four colums corresponding to the data above
```{r}
bp_df <- data.frame(patient,pre_bp,post_bp,diff_bp)
```

return only rows of the data frame where the diff bp columns is negative
```{r}
subset(patient, (diff_bp > 0))
```

Add a new column to the data frame corresponding to TRUE if the post_bp is less than 120 and FALSE if it is not
```{r}
subset(bp_df, diff_bp > 0)
```

Print the data frame out nicely in final document 
```{r}
knitr::kable(bp_df)
```

## Task 3: List Practice 

create vectors for treatment and placebo group
```{r}
treatment <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
treatment

placebo <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
```

Create a new data frame with this data in it that is similar to the data frame from task two (including the new column of TRUE/FALSE)

 Now create and store a list with two elements:
 • 1st element named treatment and contains the first data frame you created.
 • 2nd element named placebo and contains the second data frame you created.
 3. Access the first list element using three different types of syntax (the result can be a list or just the
 data frame).
 4. Use your list from question 2. In one line of code, access the pre_bp column of the placebo data
 frame